---
title: "Exploring Website Click Data"
author: "Celia Taylor"
date: "June 27, 2016"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

######Introduction
######This is from https://rpubs.com/tmcfl/simulated-click-analysis
######This tutorial is based on the EDA Exercises in the book, Doing Data Science, which I highly recommend.
######There are 31 data sets named nyt1.csv, nyt2.csv,…, nyt31.csv, which can be downloaded from GitHub.
######Each csv represents one (simulated) days worth of ads shown and clicks recorded on the New York 
######Times homepage in May 2012. Each row in the csv represents a single user.
######There are five columns: Age, Gender (0=female, 1=male), Impressions, Clicks, and 
######Signed_In status (0=not signed in, 1=signed in).

######Create a data frame called data1 from the nyt1.csv file
```{r GetData, include=TRUE, echo=TRUE}
#Create a data frame called data1 from the nyt1.csv file
fileLocation <- "http://stat.columbia.edu/~rachel/datasets/nyt1.csv"
data1 <- read.csv(url(fileLocation))
head(data1)


# Use the section below if you would rather import the data from a local file
# fileLocation <- "~/path/for/the/file/nyt1.csv"
# data1 <- read.csv("fileLocation")
```
######Take a look at the first 5 rows to get a feel for the data
```{r FeelData, include=TRUE, echo=TRUE}
#Take a look at the first 5 rows to get a feel for the data
head(data1)
```

######You can get a overview of the structure of the data by using the str() function
```{r strData, include=TRUE, echo=TRUE}
#You can get a overview of the structure of the data by using the str() function
# str stands for 'structure'
# str() returns information about each column such as name, data type, etc.
str(data1)
```
######It could also be helpful to look at a summary of the data. This gives you useful 
######information such as the number of rows in the object, and the mean, median, and 
######range (min-max) of each column

```{r summaryData, include=TRUE, echo=TRUE}
#It could also be helpful to look at a summary of the data. This gives you useful 
#information such as the number of rows in the object, and the mean, median, and 
#range (min-max) of each column.
summary(data1)
```
######Since we’re exploring, it’s helpful to visualize the distribution of some of 
######the columns, so we can see what we’re working with and if anything looks out 
######of the ordinary.
```{r DistributionAge, include=TRUE, echo=TRUE}
#Since we’re exploring, it’s helpful to visualize the distribution of some of 
#the columns, so we can see what we’re working with and if anything looks out 
#of the ordinary.
# distribution of the Age column
hist(data1$Age, main="", xlab="Age")
range(data1$Age)
```
###### Distribution of the Impressions column
```{r DistributionImpressions, include=TRUE, echo=TRUE}
# Distribution of the Impressions column
hist(data1$Impressions, main="", xlab="# of Impressions")
range(data1$Impressions)
```
###### Distribution of the Clicks column
```{r DistributionClicks, include=TRUE, echo=TRUE}
# Distribution of the Clicks column
hist(data1$Clicks, main="", xlab="# of Clicks")
range(data1$Clicks)
```
######Did you notice that there is a disproportionate number of ages recorded as zero? 
######This could skew our analysis, so we may need to account for that later.

#####Task 1
######create a new variable named Age_Group, that groups users into age categories 
######“<18”, “18-24”, “25-34”, “35-44”, “45-54”, “55-64”, and “65+”
######1.1) Cut the ages into the desired groups and add the groups to a new column 
######in data1 called Age_Group. The ‘cut’ function creates a factor with levels.
```{r CutData, include=TRUE, echo=TRUE}
######1.1) Cut the ages into the desired groups and add the groups to a new column 
######in data1 called Age_Group. The ‘cut’ function creates a factor with levels.
data1$Age_Group <- cut(data1$Age, c(-Inf, 18, 24, 34, 44, 54, 64, Inf))
# Name the levels of 'Age_Group' for readability
levels(data1$Age_Group) <- c("<18", "18-24", "25-34", "35-44", "45-54", "55-64", "65+")
```
######Take a look at the changes:
```{r HeadData, include=TRUE, echo=TRUE}
#Take a look at the changes:
head(data1)
```
#####Task 2
######For a single day, plot the distributions of ‘number of impressions’ and 
######‘click-through-rate’ by Age_Group. (CTR = clicks/impressions).
######2.1) Create a subset of data1 to exclude rows where there are no impressions 
######(if there are no impressions, we assume there will be no clicks). Name the new 
######object d1
```{r SubsetData, include=TRUE, echo=TRUE}
#2.1) Create a subset of data1 to exclude rows where there are no impressions 
#(if there are no impressions, we assume there will be no clicks). Name the new 
#object d1
d1 <- subset(data1, Impressions>0)
```
######2.2) Add a column to d1 called CTR containing the click-through-rate
```{r ClickThroughRate, include=TRUE, echo=TRUE}
#2.2) Add a column to d1 called CTR containing the click-through-rate
d1$CTR <- d1$Clicks/d1$Impressions
head(d1)
```
######2.3) Plot the distribution of Impressions>0, grouped by Age_Group, 
######using the ggplot2 package.
```{r PlotImpressions, include=TRUE, echo=TRUE}
#2.3) Plot the distribution of Impressions>0, grouped by Age_Group, 
#using the ggplot2 package.
library(ggplot2) # used for visualizations
ggplot(subset(d1, Impressions>0), aes(x=Impressions, fill=Age_Group))+
    geom_histogram(binwidth=1)
```















######2.4) Plot the distribution of CTR>0, grouped by Age_Group
```{r PlotClickThroughRate, include=TRUE, echo=TRUE}
#2.4) Plot the distribution of CTR>0, grouped by Age_Group
ggplot(subset(d1, CTR>0), aes(x=CTR, fill=Age_Group))+
    labs(title="Click-through rate by age group (05/01/2012)")+
    geom_histogram(binwidth=.025)
```



#################################################################################################

######Create a data frame called data14 from the nyt14.csv file
```{r GetData14, include=TRUE, echo=TRUE}
#Create a data frame called data14 from the nyt14.csv file
fileLocation <- "http://stat.columbia.edu/~rachel/datasets/nyt14.csv"
data14 <- read.csv(url(fileLocation))
head(data14)


# Use the section below if you would rather import the data from a local file
# fileLocation <- "~/path/for/the/file/nyt14.csv"
# data14 <- read.csv("fileLocation")
```
######Take a look at the first 5 rows to get a feel for the data
```{r FeelData14, include=TRUE, echo=TRUE}
#Take a look at the first 5 rows to get a feel for the data
head(data14)
```

######You can get a overview of the structure of the data by using the str() function
```{r strData14, include=TRUE, echo=TRUE}
#You can get a overview of the structure of the data by using the str() function
# str stands for 'structure'
# str() returns information about each column such as name, data type, etc.
str(data14)
```
######It could also be helpful to look at a summary of the data. This gives you useful 
######information such as the number of rows in the object, and the mean, median, and 
######range (min-max) of each column

```{r summaryData14, include=TRUE, echo=TRUE}
#It could also be helpful to look at a summary of the data. This gives you useful 
#information such as the number of rows in the object, and the mean, median, and 
#range (min-max) of each column.
summary(data14)
```
######Since we’re exploring, it’s helpful to visualize the distribution of some of 
######the columns, so we can see what we’re working with and if anything looks out 
######of the ordinary.
```{r DistributionAge14, include=TRUE, echo=TRUE}
#Since we’re exploring, it’s helpful to visualize the distribution of some of 
#the columns, so we can see what we’re working with and if anything looks out 
#of the ordinary.
# distribution of the Age column
hist(data14$Age, main="", xlab="Age")
range(data14$Age)
```
###### Distribution of the Impressions column
```{r DistributionImpressions14, include=TRUE, echo=TRUE}
# Distribution of the Impressions column
hist(data14$Impressions, main="", xlab="# of Impressions")
range(data14$Impressions)
```
###### Distribution of the Clicks column
```{r DistributionClicks14, include=TRUE, echo=TRUE}
# Distribution of the Clicks column
hist(data14$Clicks, main="", xlab="# of Clicks")
range(data14$Clicks)
```
######Did you notice that there is a disproportionate number of ages recorded as zero? 
######This could skew our analysis, so we may need to account for that later.

#####Task 1
######create a new variable named Age_Group, that groups users into age categories 
######“<18”, “18-24”, “25-34”, “35-44”, “45-54”, “55-64”, and “65+”
######1.1) Cut the ages into the desired groups and add the groups to a new column 
######in data14 called Age_Group. The ‘cut’ function creates a factor with levels.
```{r CutData14, include=TRUE, echo=TRUE}
######1.1) Cut the ages into the desired groups and add the groups to a new column 
######in data14 called Age_Group. The ‘cut’ function creates a factor with levels.
data14$Age_Group <- cut(data14$Age, c(-Inf, 18, 24, 34, 44, 54, 64, Inf))
# Name the levels of 'Age_Group' for readability
levels(data14$Age_Group) <- c("<18", "18-24", "25-34", "35-44", "45-54", "55-64", "65+")
```
######Take a look at the changes:
```{r HeadData14, include=TRUE, echo=TRUE}
#Take a look at the changes:
head(data14)
```
#####Task 2
######For a single day, plot the distributions of ‘number of impressions’ and 
######‘click-through-rate’ by Age_Group. (CTR = clicks/impressions).
######2.1) Create a subset of data14 to exclude rows where there are no impressions 
######(if there are no impressions, we assume there will be no clicks). Name the new 
######object d1
```{r SubsetData14, include=TRUE, echo=TRUE}
#2.1) Create a subset of data14 to exclude rows where there are no impressions 
#(if there are no impressions, we assume there will be no clicks). Name the new 
#object d14
d14 <- subset(data14, Impressions>0)
```
######2.2) Add a column to d1 called CTR containing the click-through-rate
```{r ClickThroughRate14, include=TRUE, echo=TRUE}
#2.2) Add a column to d14 called CTR containing the click-through-rate
d14$CTR <- d14$Clicks/d14$Impressions
head(d14)
```
######2.3) Plot the distribution of Impressions>0, grouped by Age_Group, 
######using the ggplot2 package.
```{r PlotImpressions14, include=TRUE, echo=TRUE}
#2.3) Plot the distribution of Impressions>0, grouped by Age_Group, 
#using the ggplot2 package.
library(ggplot2) # used for visualizations
ggplot(subset(d14, Impressions>0), aes(x=Impressions, fill=Age_Group))+
    geom_histogram(binwidth=1)
```















######2.4) Plot the distribution of CTR>0, grouped by Age_Group
```{r PlotClickThroughRate14, include=TRUE, echo=TRUE}
#2.4) Plot the distribution of CTR>0, grouped by Age_Group
ggplot(subset(d14, CTR>0), aes(x=CTR, fill=Age_Group))+
    labs(title="Click-through rate by age group (05/14/2012)")+
    geom_histogram(binwidth=.025)
```

